---
title: "Markdown"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

# Load Libraries
```{r}
library(haven)
library(dplyr)
```
 


## Load data
```{r}


allPatients = file.path("../","All_Data_Patientstudy.sav")
allDataPDGroups = file.path("../", "All_Data_PD_GroupPatientstudy.sav")

allDataPatientStudyDS = read_sav(allPatients)
allDataPDGroupsStudyDS = read_sav(allDataPDGroups)

```


## First glimpse
```{r}
glimpse(allDataPatientStudyDS)
glimpse(allDataPDGroupsStudyDS)
```
```{r}
variable.names(allDataPatientStudyDS)
```

## Codebook

```{r}
library(codebook)
library(igraph)

```
Age - age of participant in whole years
Length_Parkinson - duration of disease in years for each patient
Medication - Levodopa equivalent, rounded to nearest 50
H_Y - Hoehn Yahr
Sex - biological sex of patient
Handedness - whether the patient is right or left handed
MMSE - mini mental state examination
RAVEN - Raven's Progressive Matrices
Phonologic_Fluency - interger representation of a component a verbal fuency test
Semantic_Fluency - interger representation of a component a verbal fuency test
