---
title: "Group 4 - Parkinson's P4"
author: "Meena Mishra, Raphael Kirchgaessner, Brian Karlberg"
output: html_document
---

<span style=“color:green”>
</span>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

<span><font color="ff1493">

# Load Libraries

</span>
```{r}
library(haven)
library(dplyr)
library(tidyr)
library(ggplot2)
library(moderndive)
library(infer)
library(phia)
library(tidyverse)
```
 
<span><font color="ff1493">

# Load data

</span>

```{r}
allPatients = file.path("../","All_Data_Patientstudy.sav")
allDataPatientStudyDS = read_sav(allPatients)
tidied_dataset = read.csv("tidied-dataset.csv")
```

<span><font color="ff1493">

# EDA

</span>


```{r}
glimpse(tidied_dataset)
```

```{r}
glimpse(allDataPatientStudyDS)
```
```{r include=FALSE}
variable.names(allDataPatientStudyDS)
```

<span><font color="ff1493"> 

## Relabeled columns

</span>

```{r}
allDataPatientStudyDS %>%
mutate(Sex = factor(Sex, labels = c("Female", "Male")),Group = factor(Group, labels = c("PD", "Control")))
```
```{r}
tidied_dataset <- tidied_dataset %>%
mutate(Sex = factor(Sex, labels = c("Male", "Female")),Group = factor(Group, labels = c("Control", "PD")))
```

<span><font color="ff1493">
# Replication section
</span>

<span><font color="ff1493"> 
## Table 1
We were able to create table 1 as it appears in the paper, getting al the same values for means, range and standard deviation of age, medication, length of medication and parkinson stage (H-Y)
</span>

```{r}
#Table 1 recreated using allDataPatientStudyDS dataframe
Table1 <- tidied_dataset %>%
  filter(Group == "PD")%>%
select("Group","Age", "Medication", "Length_Parkinson","H_Y")
Table1
skimr::skim(Table1)
```

<span><font color="ff1493"> 
## Table 3
Summary statistics for all control variables were replicated and they matched th evalues in the paper.
</span>

```{r}
#Table 3 using tidied_dataset
Table3 <- tidied_dataset %>%
  select("Age", "Sex", "Handedness","MMSE","RAVEN","Group")%>%
  group_by(Group)%>%
  summarize(mean_age = mean(Age), mean_handedness = mean(Handedness),mean_MMSE = mean(MMSE), mean_Raven = mean(RAVEN), count = n())
Table3_count_gender <- tidied_dataset %>%
  select("Sex","Group")%>%
  group_by(Group,Sex)%>%
  summarize(count=n())
t(Table3)
Table3_count_gender
```
<span><font color="ff1493"> 
## Table 4
We were able to replicate table 4 as it appears in the paper. Just to point out that COWAT is a combined term for phonologic fluency and semantic fluency. So we have replicated the numbers for COWAT without calling it out. 
</span>

```{r}
#Table 4 using tidied dataset
Table4 <- tidied_dataset %>%
  group_by(Group)%>%
select("Phonologic_Fluency","Semantic_Fluency","BNT", "SyllSec_Norm", "SyllSec_Ruis", "NPro_Dur_Norm","NPro_Dur_Ruis")%>%
  summarise(mean_Phonologic_Fluency = mean(Phonologic_Fluency), sd_phonologic_fluency = sd(Phonologic_Fluency),
            mean_Semantic_Fluency = mean(Semantic_Fluency), sd_Semantic_Fluency = sd(Semantic_Fluency),
            mean_BNT = mean(BNT), sd_BNT = sd(BNT),
            mean_Syllsec_Norm = mean(SyllSec_Norm), sd_Syllsec_Norm = sd(SyllSec_Norm),
            mean_Syllsec_Ruis = mean(SyllSec_Ruis), sd_SyllSec_Ruis = sd(SyllSec_Ruis),
            mean_NPro_Dur_Norm = mean(NPro_Dur_Norm), sd_NPro_Dur_Norm = sd(NPro_Dur_Norm),
            mean_NPro_Dur_Ruis = mean(NPro_Dur_Ruis), sd_NPro_Dur_Ruis = sd(NPro_Dur_Ruis))
Table4
t(Table4)
```
<span><font color="ff1493"> 
## Table 9
Table 9 was replicated without issues.
</span>

```{r}
#Table 9 using using tidied dataset
Table9_raw <- tidied_dataset %>%
  select("Group","Nrm_Reformulation","Nrm_FilledPause","Nrm_Repetition","Nois_Reformulation","Nois_FilledPause", "Nois_Repetition")
Table9 <- Table9_raw %>%
  gather (key = "Error_type", value = "Total_Number",-Group,"Nrm_Reformulation","Nrm_FilledPause","Nrm_Repetition","Nois_Reformulation","Nois_FilledPause", "Nois_Repetition")
Table9
aggregate(Total_Number~Error_type+Group,Table9,sum)
 
```
<span><font color="ff1493"> 
## Figure 1 and Figure 2
Replicating figure 1 and Fig 2 was hard. We found that the values for "percentage of semantic error repaired", provided in the raw dataset were not actually percentages. Closer analysis revealed that the values for "percentage semantic error repaired " may have been fractions. We recalculated percentage of semantic error repaired and used those values to come up with fig 1 and fig2. 
</span>

```{r }
#Replicating figure 1 using r=the values from raw dataset
Fig1_raw <- allDataPatientStudyDS %>%
  select("Group","Pct_NormRep","Pct_NoiseRep","PctSemNoiseRep","PctSemNormRep")%>%
  filter(Group == 0,complete.cases(Pct_NormRep,Pct_NoiseRep,PctSemNoiseRep,PctSemNormRep))%>%
  summarise(mean_Pct_NormRep=mean(Pct_NormRep),mean_Pct_NoiseRep=mean(Pct_NoiseRep),mean_PctSemNoiseRep=mean(PctSemNoiseRep),mean_PctSemNormRep=mean(PctSemNormRep, count=n()))
 
Fig1_table <- Fig1_raw %>%
  gather (key = "Error", value = "Percentage_Repaired","mean_PctSemNormRep","mean_PctSemNoiseRep","mean_Pct_NormRep","mean_Pct_NoiseRep")
ggplot(data = Fig1_table) +
  geom_col(mapping = aes(x = Error,y=Percentage_Repaired), binwidth = 0.5)
 
  
```
```{r}
#Replicating figure 2 using r=the values from raw dataset
Fig2_raw <- allDataPatientStudyDS %>%
  select("Group","Pct_NormRep","Pct_NoiseRep","PctSemNoiseRep","PctSemNormRep")%>%
  filter(complete.cases(Pct_NormRep,Pct_NoiseRep,PctSemNoiseRep,PctSemNormRep))%>%
  group_by(Group)%>%
  summarise(mean_Pct_NormRep=mean(Pct_NormRep),mean_Pct_NoiseRep=mean(Pct_NoiseRep),mean_PctSemNoiseRep=mean(PctSemNoiseRep),mean_PctSemNormRep=mean(PctSemNormRep, count=n()))
 
Fig2_table <- Fig2_raw %>%
  gather (key = "Error", value = "Percentage_Repaired","mean_PctSemNormRep","mean_PctSemNoiseRep","mean_Pct_NormRep","mean_Pct_NoiseRep")
Fig2_table %>%
  ggplot() +
  geom_col(mapping = aes(x = Error,y=Percentage_Repaired), binwidth = 0.5)+
  facet_wrap(~Group)
  
```
<span><font color="ff1493">
### Recalculating the values for percentage of semantic error and using that for replicating fig 1 and fig 2. The figues still do not resemble with what is in the paper. 
</span>
```{r}
Fig_table <- allDataPatientStudyDS %>%
  group_by(Group)%>%
  select("SemErrors_Norm","Nrm_Sem_Rep", "SemErrors_Noise","Nois_Sem_Rep", "Norm_ErrorsRepaired","Noise_ErrorsRepaired","TotalErrorsNorm_N","TotalErrorsNois_N" )%>%
  mutate(Nrm_Perc_Sem_Error_Repaired = Nrm_Sem_Rep/SemErrors_Norm * 100, Nois_Per_Sem_Error_Repaired = Nois_Sem_Rep/SemErrors_Noise *100, Nrm_Perc_Total_Error_Rep = Norm_ErrorsRepaired/TotalErrorsNorm_N *100, Nois_Per_Total_Error_Rep = Noise_ErrorsRepaired/TotalErrorsNois_N *100)%>%
replace(is.na(.), 0)%>%
group_by(Group) %>% 
summarize(Mean_Nrm_Perc_Sem_Error_Repaired = mean(Nrm_Perc_Sem_Error_Repaired),Mean_Nois_Per_Sem_Error_Repaired = mean(Nois_Per_Sem_Error_Repaired), Mean_Nrm_Perc_Total_Error_Repaired = mean(Nrm_Perc_Total_Error_Rep),Mean_Nois_Perc_Total_Error_Repaired = mean(Nois_Per_Total_Error_Rep))%>%
rename(Sem_Normal=Mean_Nrm_Perc_Sem_Error_Repaired, Sem_Noise=Mean_Nois_Per_Sem_Error_Repaired, Tot_Normal=Mean_Nrm_Perc_Total_Error_Repaired, Tot_Noise=Mean_Nois_Perc_Total_Error_Repaired)
Fig_table
skimr::skim(Fig_table)
```
```{r}
Fig1 <- Fig_table  %>%
  filter(Group==0)%>%
  gather (key = "Error", value = "Mean_Percentage_Repaired","Sem_Normal","Sem_Noise","Tot_Normal","Tot_Noise" )%>%
  ggplot() +
  geom_col(mapping = aes(x = Error,y=Mean_Percentage_Repaired), binwidth = 0.5)
 
Fig1
```


```{r}
Fig2 <- Fig_table  %>%
  gather (key = "Error", value = "Mean_Percentage_Repaired","Sem_Normal","Sem_Noise","Tot_Normal","Tot_Noise" )%>%
  ggplot() +
  geom_col(mapping = aes(x = Error,y=Mean_Percentage_Repaired), binwidth = 0.5)+
  facet_wrap(~Group)
Fig2
```
<span><font color="ff1493">
# Table 11: Regression analysis of Parkinson's disease measures
Simple linear refressions were run for task performances as explained by Parkinson's disease measures. Duration of PD, Medication and H_Y score were used as disease measures. 

Since there are a few NA vales for H-Y variales, we used only the complete case of H_Y for our analysis. Table 11 reports just the significant findings. 



```{r}
Regression_table <- tidied_dataset %>%
  filter(Group == "PD",complete.cases(H_Y) )
```

```{r}
Model_sem_fluency <- lm(Semantic_Fluency~H_Y,Regression_table)
get_regression_table(Model_sem_fluency)
broom::glance(Model_sem_fluency)
```

```{r}
Model_BNT <- lm(BNT~H_Y,Regression_table)
get_regression_table(Model_BNT)
broom::glance(Model_BNT)
```
```{r}
Model_ProductionSpeedNormal <- lm(SyllSec_Norm~H_Y,Regression_table)
get_regression_table(Model_ProductionSpeedNormal)
broom::glance(Model_ProductionSpeedNormal)
```
```{r}
Model_ProductionSpeedNoise <- lm(SyllSec_Ruis~H_Y,Regression_table)
get_regression_table(Model_ProductionSpeedNoise)
broom::glance(Model_ProductionSpeedNoise)
```
```{r}
Model_DurationNorm <- 
lm(NPro_Dur_Norm~H_Y,Regression_table)
get_regression_table(Model_DurationNorm)
broom::glance(Model_DurationNorm)
```
```{r}
Model_DurationNoise <- 
lm(NPro_Dur_Ruis~H_Y,Regression_table)
get_regression_table(Model_DurationNoise)
broom::glance(Model_DurationNoise)
```
```{r}
Model_PectSemErr_Rep <- 
lm(PctSemNormRep~H_Y,allDataPatientStudyDS)
get_regression_table(Model_PectSemErr_Rep)
broom::glance(Model_PectSemErr_Rep)
```
```{r}
Model_Ntw_Perc_Sem_NErrors <- 
lm(Ntw_Perc_Sem_NErrors~H_Y,allDataPatientStudyDS)
get_regression_table(Model_Ntw_Perc_Sem_NErrors)
broom::glance(Model_Ntw_Perc_Sem_NErrors)
```
# t test to test if the difference in the mean BNT score between the control and PD group is not equal to 0.
Inspite of non-normal distribution, a two-sided t test for unequal variance (Welch's two sample t test) was done. With a p value of 0.494 there is not enough evidence to support the alternate hypothesis. In other words we fail to reject the Null hypothesis that difference between the mean BNT score for the two groups is equal to 0. 
```{r}
tidied_dataset%>%
  ggplot(aes(x=BNT)) +
  geom_histogram()+
  facet_wrap(~Group)
  
  
```

```{r}
t.test(BNT ~ Group, data = tidied_dataset, null = 0, 
       alternative = "two.sided",var.equal = TRUE)
```
```{r}
t.test(Phonologic_Fluency ~ Group, data = tidied_dataset, null = 0, 
       alternative = "two.sided",var.equal = TRUE)
```

```{r}
t.test(Semantic_Fluency ~ Group, data = tidied_dataset, null = 0, 
       alternative = "two.sided",var.equal = TRUE)
```
```{r}
library("car")
qqPlot(BNT~Group,tidied_dataset)
```
```{r}
library("car")
qqPlot(Phonologic_Fluency~Group,tidied_dataset)
```
```{r}
library("car")
qqPlot(Semantic_Fluency~Group,tidied_dataset)
```

```{r}
lmMod <- lm(BNT ~ Group, data=tidied_dataset)
par(mfrow=c(2,2)) # init 4 charts in 1 panel
plot(lmMod)
```
```{r}
lmMod_Semantic <- lm(Semantic_Fluency ~ Group, data=tidied_dataset)
lmMod_Semantic_aug <- broom::augment(lmMod_Semantic)%>%
  ggplot(aes(x = .fitted, y = .resid)) +
  geom_point()
lmMod_Semantic_aug
```
```{r}
lmMod_BNT <- lm(BNT ~ Group, data=tidied_dataset)
lmMod_BNT_aug <- broom::augment(lmMod_BNT)%>%
  ggplot(aes(x = .fitted, y = .resid)) +
  geom_point()
lmMod_BNT_aug
```
```{r}
lmMod_Phonologic_Fluency <- lm(Phonologic_Fluency ~ Group, data=tidied_dataset)
lmMod_Phonologic_Fluency_aug <- broom::augment(lmMod_Phonologic_Fluency)%>%
  ggplot(aes(x = .fitted, y = .resid)) +
  geom_point()
lmMod_Phonologic_Fluency_aug
```
```{r}

Error_type <- read_csv("Error_type.csv")

```
```{r}
Error_type <- Error_type%>%
  mutate(Group = factor(Group), Noise = factor(Noise))
```

<ul style="color:deeppink"> 
## ANOVA Results for Disfluencies

#### For all disfluencies we found no significant results for noise, interaction between noise and group. For reformulation, we do see a significant effect on disfluency based on group. Our results here do not match the paper. For the control group it seems like the total disfluencies is slightly but not significantly more under noise condition. However the paper finds that there is a significant interaction of condition and group.
</ul>

```{r}
model_Disfluencies <- lm(Disfluencies ~ Group + Noise, data=Error_type)
anova (model_Disfluencies)
moderndive::get_regression_table(lm(Disfluencies ~ Group + Noise, data=Error_type))
```

```{r}
model_Disfluencies_interact <- lm(Disfluencies ~ Group * Noise, data=Error_type, 
        contrasts = list(Group = contr.sum, 
                         Noise = contr.sum))
Anova(model_Disfluencies_interact, type = 2)
```

```{r}
plot(interactionMeans(model_Disfluencies_interact))
```

```{r}
model_Reformulation <- lm(Reformulation ~ Group + Noise, data=Error_type)
anova (model_Reformulation)
moderndive::get_regression_table(lm(Reformulation ~ Group + Noise, data=Error_type))
```
```{r}
model_Reformulation_interact <- lm(Reformulation ~ Group * Noise, data=Error_type, 
        contrasts = list(Group = contr.sum, 
                         Noise = contr.sum))
Anova(model_Reformulation_interact, type = 2)
```

```{r}
plot(interactionMeans(model_Reformulation_interact))
```

```{r}
model_Filled_Pause <- lm(Filled_Pause ~ Group + Noise, data=Error_type)
anova (model_Filled_Pause)
moderndive::get_regression_table(lm(Reformulation ~ Group + Noise, data=Error_type))
```
```{r}
model_Filled_Pause_interact <- lm(Filled_Pause ~ Group * Noise, data=Error_type, 
        contrasts = list(Group = contr.sum, 
                         Noise = contr.sum))
Anova(model_Filled_Pause_interact, type = 2)
```
```{r}
plot(interactionMeans(model_Filled_Pause_interact))
```

```{r}
model_Repetition <- lm(Repetition ~ Group + Noise, data=Error_type)
anova (model_Repetition)
moderndive::get_regression_table(lm(Repetition ~ Group + Noise, data=Error_type))
```
```{r}
model_Repetition_interact <- lm(Repetition ~ Group * Noise, data=Error_type, 
        contrasts = list(Group = contr.sum, 
                         Noise = contr.sum))
Anova(model_Repetition_interact, type = 2)
```
```{r}
plot(interactionMeans(model_Repetition_interact))
```

```{r}
model_Disfluencies <- lm(Disfluencies ~ Group + Noise, data=Error_type)
anova (model_Disfluencies)
moderndive::get_regression_table(lm(Disfluencies ~ Group + Noise, data=Error_type))
```
```{r}
model_Disfluencies_Interaction <- lm( Disfluencies ~ Group * Noise, data=Error_type, 
        contrasts = list(Group = contr.sum, 
                         Noise = contr.sum))
Anova(model_Disfluencies_Interaction, type = 2)
```
```{r}
plot(interactionMeans(model_Disfluencies_Interaction))
```




















