---
title: "P2-Addition"
author: "Meena Mishra, Raphael Kirchgaessner, Brian Karlberg"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```
## 1. Following best practice for collaboration and teanwork

We wanted to keep track of changes we make to the documents for our project. Our team decided to use Github to enable collaboration and best practices for data and project sharing. We looked at the data sets available to us for our project at https://osf.io/37u7u/. We followed the guidelines suggested by Ellis & Leek paper.

## 2. Load Libraries
```{r}
library(haven)
library(dplyr)
```

# 3. Raw data
The data available to us was not at its rawest form. It had already been tidied and some data excluded from the data set because of greater variance in the test results. This information is shared by the authors in the "Result" section of the paper-"To compare reaction times, incorrect trials and trials that deviated more than 2 SD from the mean per condition were removed from the data set. Per task the data loss is reported." Our dataset also came with no codebook. 

# 4. Load data 
The first step was to load and save the available data sets. The data files were available in SPSS file format with .sav extentions. We saved these files as tibble dataframes. There were 2 data files available for our project.
```{r}

allPatients = file.path("All_Data_Patientstudy.sav")
allDataPDGroups = file.path("All_Data_PD_GroupPatientstudy.sav")

allDataPatientStudyDS = read_sav(allPatients)
allDataPDGroupsStudyDS = read_sav(allDataPDGroups)

```

# 5. First glimpse 
Next we glimpsed at our datasets. There were 34 observations and 90 variables in "allDataPatientStudyDS" where as the file titled "allDataPDGroupsStudyDS" had 18 observations and 92 variables. We dug in deeper as to the kinds of variables available in both datasets. Closer examinations revealed that the file titled "allDataPatientStudyDS" included data on all 34 subjects in the study and would be sufficient for the purpose of our project. 

```{r ECHO=FALSE}
glimpse(allDataPatientStudyDS)
glimpse(allDataPDGroupsStudyDS)
```
## 6. Looking at all variable names in "allDataPatientStudyDS"

```{r}
variable.names(allDataPatientStudyDS)

```

## 7. Tidying Dataset

Per Tidy guidelines (Ellis & Leek paper):
1. Each variable in our data table is in a seperate column.
2. Each observation corresponds to each of the study subject and is on a different row.
3. However, all data for all different kinds of Parkinson tests employed in the study are in the same single data tabe. 

We selected all columns/variables that were needed to recreate tables and graphs for our project by using the "select" function and saved the new dataset as "patientStudy". Next We glimpsed at "patientStudy" table. 

```{r}

patientStudy <- allDataPatientStudyDS %>%
  select("Subject","Age", "Length_Parkinson", "Medication", "H_Y","Sex", "Group", "Handedness", "MMSE","RAVEN", "Phonologic_Fluency", "Semantic_Fluency", "BNT", "SyllSec_Norm", "SyllSec_Ruis", "NPro_Dur_Norm", "NPro_Dur_Ruis", "Nrm_Reformulation", "Nrm_FilledPause", "Nrm_Repetition", "DisfluenciesNorm", "Nois_Reformulation", "Nois_FilledPause", "Nois_Repetition", "DisfluenciesNoise", "Nrm_Sem_Unrep", "Nrm_Sem_Rep", "SemErrors_Norm", "Nrm_Phon_Unrep", "Nrm_Phon_Rep", "PhonErrors_Norm", "Nrm_Gram_Unrep", "Nrm_Gram_Rep", "GramErrors_Norm", "Norm_ErrorsUnrepaired", "Norm_ErrorsRepaired", "TotalErrorsNorm_N", "Nois_Sem_Unrep", "Nois_Sem_Rep", "SemErrors_Noise", "Nois_Phon_Unrep", "Nois_Phon_Rep", "PhonErrors_Nois", "Nois_Gram_Unrep", "Nois_Gram_Rep", "GramErrors_Nois", "Noise_ErrorsUnrepaired", "Noise_ErrorsRepaired", "TotalErrorsNois_N", "Norm_ErrorsRepaired_ExDisf", "TotalErrorsNormal_exDisf", "Noise_ErrorsRepaired_ExDisf", "TotalErrorsNoise_exDisf")

```
```{r ECHO=FALSE}

glimpse(patientStudy)
```

## 8. Converting data types for variables 
 We looked at the data type and structure for all the varables in out final table. We discovered that the variable "Group" was stored as type "double" in the patientStudy table. In order to regenerate the graphs and tables by "Group" (control and Parkinson), we decided to convert the variable "Group" from double to a factor. We were curious about the differences in the test results between control and Parkinson group, with respect to gender, therefore we converted the variable "Sex" from "character to factor as well. 
 
```{r Converting}
patientStudy$Sex <- factor(patientStudy$Sex)
patientStudy$Group <- factor(patientStudy$Group) 
```

## 9. Saving the final dataset for our project as tidied dataset - tidied-dataset
Finally we converted the final dataset for our project as a .csv file. 

```{r Write csv}
write.csv(patientStudy, "tidied-dataset.csv")
```



