---
title: "P2-Addition"
author: "Meena Mishra, Raphael Kirchgaessner, Brian Karlberg"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

# Load Libraries
```{r}
library(haven)
library(dplyr)
```
 


## Load data

Both datasets are located in the exact some path as the P2-Addition.rmd file.

```{r}

allPatients = file.path("All_Data_Patientstudy.sav")
allDataPDGroups = file.path("All_Data_PD_GroupPatientstudy.sav")

allDataPatientStudyDS = read_sav(allPatients)
allDataPDGroupsStudyDS = read_sav(allDataPDGroups)

```


## First glimpse
```{r}
glimpse(allDataPatientStudyDS)
glimpse(allDataPDGroupsStudyDS)
```
```{r}
variable.names(allDataPatientStudyDS)
```

## Tidying Dataset

We evaluted that we only need one file to recreate important tables and figures. That is the file "All_Data_Patientstudy.sav". 
To generate a new dataset, we selected all importants columns from the existing dataset and had a second glimpse.

```{r}

patientStudy <- allDataPatientStudyDS %>%
  select("Subject","Age", "Length_Parkinson", "Medication", "H_Y","Sex", "Group", "Handedness", "MMSE","RAVEN", "Phonologic_Fluency", "Semantic_Fluency", "BNT", "SyllSec_Norm", "SyllSec_Ruis", "NPro_Dur_Norm", "NPro_Dur_Ruis", "Nrm_Reformulation", "Nrm_FilledPause", "Nrm_Repetition", "DisfluenciesNorm", "Nois_Reformulation", "Nois_FilledPause", "Nois_Repetition", "DisfluenciesNoise", "Nrm_Sem_Unrep", "Nrm_Sem_Rep", "SemErrors_Norm", "Nrm_Phon_Unrep", "Nrm_Phon_Rep", "PhonErrors_Norm", "Nrm_Gram_Unrep", "Nrm_Gram_Rep", "GramErrors_Norm", "Norm_ErrorsUnrepaired", "Norm_ErrorsRepaired", "TotalErrorsNorm_N", "Nois_Sem_Unrep", "Nois_Sem_Rep", "SemErrors_Noise", "Nois_Phon_Unrep", "Nois_Phon_Rep", "PhonErrors_Nois", "Nois_Gram_Unrep", "Nois_Gram_Rep", "GramErrors_Nois", "Noise_ErrorsUnrepaired", "Noise_ErrorsRepaired", "TotalErrorsNois_N", "Norm_ErrorsRepaired_ExDisf", "TotalErrorsNormal_exDisf", "Noise_ErrorsRepaired_ExDisf", "TotalErrorsNoise_exDisf")

glimpse(patientStudy)
```

### Converting variables

In order to follow best practices dataset tidying, some columns should be converted from numeric to factor.
In our case this should be done for the columns Sex and Group.

```{r Converting}
patientStudy$Sex <- factor(patientStudy$Sex)
patientStudy$Group <- factor(patientStudy$Group) 
```


### Write CSV

The tidyed dataset will be created in the same folder.

```{r Write csv}
write.csv(patientStudy, "tidied-dataset.csv")
```



